---
title: "Github Notes"
author: "Rohan Parmar"
date: "2023-02-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Index:
## 1: Core Concepts - Page 2
## 2: Basics - Foundational Commands - Page 3
## 3: Advanced Git - Page 7
## 4: Comparing Differences - Page
## 5: Social Coding - Page
## 6: Github Issues - Page
## 7: Github Gists - Page

\newpage
# 1 - Core Concepts:

**Git**: Decentralised / Distributed Control System.

* Can still use a central server
* Scales well -> Created for Linux Kernel Project (Linus Torvalds)
* Local operations -> Super fast
* Open Source
* Active Community

**Repository**: Collection of files managed by Git.

* Files/ structure lives in **.git** folder

**Commits and Files**:

* Updates are saved in snapshots called *Commits*
* Git commits *files* **not** *folders*
* Use *dummy files* to "save" empty folders
* Commits are saved on a timeline, known as *the branch*
* At least one branch in a repo, called "master"
* Can create additional branches!

**Github**: Git hosting repository service provided by Github, Inc

* Most popular git hosting platform
* Comprehensive, strong features
* Claim to fame: Free unlimited public repositories
* Many open source projects!
* Extra features: issue tracking, web pages, etc.

\newpage

# 2 - Basics - Foundational Commands:

## Getting Started / New Repo:

Creating a new git repository locally in terminal:

```{R, eval = FALSE}
# Print working directory
pwd
# List all files including hidden ones, and in long format
ls -al
# Create new directory -> We will use this as our base folder
mkdir projects
# Move into the newly created directory
cd projects/
# Create new repository within the projects folder that will be our current project folder
git init demo
# Move into the demo folder
cd demo/
  
# At this point, the terminal should have a (master) tag on the command line, indicating 
# that you are on the default branch on the git repo, as follows:
(master) demo $
```

## Git States:

**Working Directory**: All the files and folders for your application, which may or may not be managed by Git. Git is however, aware of those files.

**Staging Area**: Used to prepare for the next commit. Files are moved from the working directory to here, to the repository.

**Repository**: All the committed or saved changes to the git repository, stored in the ".git" folder. Everything here is part of Git's history.

**Remote**: 4th stage of the Git State, it is just the local repository saved on a remote server.

## First Commit:

To check out what is happening on the repository, you can use:

```{r, eval=FALSE}
git status
```

It indicates which branch you are in, what the current state of the repo is.

Now, we should create a READ ME file, which is a Markdown text file used to indicate what the repository is all about.

```{r, eval=FALSE}
mate README.md
```

The *git status* command will now indicate that we have an untracked file in the repo.

To add the file from the Working Directory to the Staging Area:

```{r, eval=FALSE}
git add README.md
```

The *git status* command will indicate that we now have changes to be committed.

Committing the staged file:

```{r, eval=FALSE}
git commit -m "First file in demo repo"
```

The *git status* will now indicate we have nothing to commit.

## Starting with an Existing Project:

Process:

* cd to the existing project folder
* use git init with the current folder
* add files to the repo
* commit files

```{r, eval=FALSE}
cd projects/
git init .
git status
git add .
git commit -m "Initialising the repository"
```

## Committing Details with Log and Show:

Git log displays the commit history.

```{r, eval=FALSE}
(master) demo $ git log
```

Output:

```{r, eval=FALSE}
commit 2g2vjgv4hgc22456hgvhfc23434hgv5jgv1jhvb5
Author: Rohan Parmar <rohanparmar@git>
Date: Sat Aug 15 00:16:12 2023 -400

  Initialising the repository
```

Anatomy of the log:

Commit ID: used to uniquely identify commits in the repo \newline
Author: Person who made the commit \newline
Date: Date and Time of commit \newline
Message: The commit message

\textcolor{red}{We can get the same information with more details about the diff using } "git show"

\newpage
## Express Commit - Combining Commands:

To see which files git is tracking:

```{r, eval=FALSE}
(master) demo $ git ls-files
```

Add a new file:

```{r, eval=FALSE}
(master) demo $ touch new.file
```

The new file is not tracked, as proved by git status.

Remove the file:

```{r, eval=FALSE}
(master) demo $ rm new.file
```

The express commit relies on tracked files to work.

```{r, eval=FALSE}
(master) demo $ git commit -am "Updating Readme"
```

This stages all changes immediately and commits it.

## Backing out changes - Unstaging:

```{r, eval=FALSE}
(master) demo $ git reset HEAD README.md
```

To discard the changes entirely, revert back to last state in the repo

```{r, eval=FALSE}
(master) demo $ git checkout --README.md
```

*Sometimes, no news is good news*

## History and Alias:

```{r, eval=FALSE}
(master) demo $ git log --oneline --graph --decorate --all
```

We can configure a new command for this to make it easier to call

```{r, eval=FALSE}
(master) demo $ git config --global alias.hist "log --oneline --graph --decorate --all" 
```

Checking if the entry made it in:

```{r, eval=FALSE}
(master) demo $ git config --global --list
```

To use the new command:

```{r, eval=FALSE}
(master) demo $ git hist
```

Can also pass files!:

```{r, eval=FALSE}
(master) demo $ git hist --README.md
```

## Rename and Delete files:

Rename:

```{r, eval=FALSE}
(master) demo $ git mv example.txt demo.txt
```

Delete:

```{r, eval=FALSE}
(master) demo $ git rm demo.txt
```

## Exluding Unwanted Files:

In the command line:

```{r, eval=FALSE}
(master) demo $ mate .gitignore
```

In the file that opens up, write the name of the files you want ignored

```{r, eval=FALSE}
*.log # all log files
application.log # just this one
```

\newpage

# 3 - Advanced Git:

## Comparing Differences:

Compare what has changed.

```{r, eval=FALSE}
(master) demo $ git diff {commit ID} {commit ID}
```

## Branching and Merge Types:

Branch = Timeline of Commits.

More accurately, branches are labels we give to commits. Deletion removes label only.

Types of merges:

* Fast Forward Merge: Simplest case, no additional work has been detected on the parent branch. Can be disabled.
* Automatic Merge: Happens when Non-Conflicting Merge is detected, preserves both timelines and merges commit on destination.
* Manual Merge: Automatic merge not possible due to conflicting merge state. Conflicts must be resolved automatically before moving forward.

## Special Markers:

Just like "pointers".

HEAD -> Points to Last Commit of Current Branch, and can be moved.

## Simple Branching Example:

Creating a new branch and switching to it:

```{r, eval=FALSE}
(master) demo $ git checkout -b {branch name}
```

To check the difference between the branches:

```{r, eval=FALSE}
({branch name}) demo $ git diff {branch name} master
```

To integrate the changes on the branch:

* Switch to parent branch
* Merge with previous branch
* Delete previous branch

```{r, eval=FALSE}
({branch name}) demo $ git checkout master
(master) demo $ git merge {branch name}
(master) demo $ git branch -d {branch name}
```
